Summary of the Iberian Bees Database

### 2022-03-01 Manuscript/Summary/Summary.Rmd
minor edits

```{r include=FALSE}
library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz",  header=T, quote="\"", sep=",")

master <- read.csv("../../Data/Processing_iberian_bees_raw/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
colnames(data)
colnames(master)
data[which(!data$accepted_name %in% master$accepted_name),] #OK
master[which(!master$accepted_name %in% data$accepted_name),] 
#we are missing ~100 species, mostly parasites!
length(unique(data$accepted_name))
nrow(master)
```

@@

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#map
library(rgdal)
library(ggplot2)

@@

plotmap <- ggplot(data = data, aes(x = Longitude, y = Latitude)) +  # plot the data
           geom_polygon(data = ESPTmap, aes(x = long, y = lat, group = group),            # plot Spain
                        color = "black", fill = "white") + coord_fixed() +    # coord_fixed() ensures that one unit on the x-axis is the same length as one unit on the y-axis
           geom_point(color = "dodgerblue4", size = 2, shape = ".") +                   # graphical parameters for points
           stat_density2d(aes(x = jitter(Longitude),                           # create the density layer based on where the points are
                              y = jitter(Latitude),  fill = ..level.., alpha = ..level..),   # colour and transparency depend on density
           geom = "polygon", colour = "grey95",size=0.3) +            # graphical parameters for the density layer
           scale_fill_gradient(low = "yellow", high = "red") +                 # set colour palette for density layer
           scale_alpha(range = c(.25, .5), guide = FALSE) +                    # set transparency for the density layer 
           theme(axis.title.x=element_blank(), axis.text.x=element_blank(),  # don't display x and y axes labels, titles and tickmarks 
           axis.ticks.x=element_blank(),axis.title.y=element_blank(),   
           axis.text.y=element_blank(), axis.ticks.y=element_blank(),
           #text=element_text(size=18),legend.position = c(.9, .15),       # size of text and position of the legend
           panel.grid.major = element_blank(),                            # eliminates grid lines from background
           panel.background = element_blank())                            # set white background
# now plot, it takes a while!
plotmap
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width="70%"}
#code courtesy of Paco Rodriguez.
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
crs.geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs.laea <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz",  header=T, quote="\"", sep=",")

#Prepare data
data$Longitude <- as.numeric(data$Longitude)
data$Latitude <- as.numeric(data$Latitude)
data <- data %>% filter(!is.na(Latitude) & !is.na(Longitude))
#Keep points on balearic islands
data_1 <- data %>% filter(Province=="Balears, Illes")
data_1$Country_1 <- NA

#Filter points in Spain and Portugal
data$Country_1 <- NA
data$Country_1 <- maps::map.where(x = data$Longitude, y = data$Latitude)
data <- data %>% filter(Country_1=="Spain" | Country_1=="Portugal")
#Rbind points that are in Spain with the ones of Balearic islands
dat <- rbind(data, data_1)

datos.sf <- st_as_sf(dat,
                     coords = c("Longitude", "Latitude"),
                     crs = crs.geo)

datos.laea <- st_transform(datos.sf,
                           crs = crs.laea)

@@

ras.counts <- rasterize(datos.laea[, "accepted_name"], ras,
               field = "accepted_name", fun = "count")
#str(datos.laea)
datos.laea$total <- datos.laea$Male + datos.laea$Female + datos.laea$Worker + datos.laea$Not.specified 
ras.counts2 <- rasterize(datos.laea[, "total"], ras,
               field = "total", fun = "sum")
ras2 <- raster(datos.laea, resolution = 30000)
ras.richness <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {length(unique(na.omit(x)))} )
library(vegan)
ras.chao <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {estimateR(t(as.matrix(x = table(x), nrow = 1, ncol = x)))[2,1]} ) #2 chao, 4 ACE #TO DO THIS WELL, we need to recover abundances.

#levelplot(log(ras.counts+1), margin = FALSE, scales = list(draw = FALSE),
 #         par.settings = viridisTheme())
levelplot(log(ras.counts2+1), margin = FALSE, scales = list(draw = FALSE),
          par.settings = viridisTheme())
#levelplot(log(ras.richness+1), margin = FALSE, scales = list(draw = FALSE),
 #         par.settings = viridisTheme())
#levelplot(ras.chao, margin = FALSE, scales = list(draw = FALSE),
 #         par.settings = viridisTheme())

@@

```{r echo=FALSE, fig.align='center', out.width="70%"}
#Prepare top 20 records
top_20 <- data %>% filter(!accepted_name=="Apis mellifera") %>% group_by(accepted_name) %>%
summarise(no_rows = length(accepted_name)) %>% 
arrange(-no_rows) %>% slice(1:20)
#Fix row position
top_20$accepted_name <- factor(top_20$accepted_name, levels = top_20$accepted_name)
#Plot
ggplot(top_20, aes(x=accepted_name, y=no_rows)) + 
geom_bar(stat = "identity", fill="#287D8EFF") +
theme_classic()+
theme(axis.text.x=element_text(angle=45, hjust=1, size=8, face = "italic")) +
ylab("Number of records") + 
xlab("Species") + 
scale_y_continuous(expand = c(0, 0))

@@

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Prepare top 20 records
by_year <- data %>% filter(!accepted_name=="Apis mellifera")

p1 <- ggplot(by_year, aes(Year)) +
  geom_histogram(fill="#287D8EFF") + theme_classic() + ylab("Number of records")+
  scale_y_continuous(expand = c(0, 0))

@@

month_sum$month_name <- factor(month_sum$month_name, levels =  c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec"))

p2 <- ggplot(data=month_sum, aes(x=month_name, y=no_rows)) +
  geom_bar(stat="identity", fill="#287D8EFF") + theme_classic()+
  scale_y_continuous(expand = c(0, 0)) +
  ylab("Number of records") + xlab("Months")

@@

p1 / p2

@@

#table of
#Name	
#Contact	
#Number of records as main dataholder	
#Number of records as co-dataholder
#email?? Maually :(
#default <- par()$mar
#par(mar = c(13.1, 4.1, 4.1, 2.1))
#barplot(head(sort(table(data2$Authors.to.give.credit), decreasing = TRUE),20), las = 2, cex.names = 0.7)
#par(mar = default)
#Can be done prettier.
#Ideally, we can merge all authors, split and count individually.

@@

#sort(unique(author_list))[-1] #79 contributors!
#write.csv(sort(unique(author_list))[-1], "data/authors.csv")

@@

"Alejandro Núñez"
,"Amparo Lázaro"
,"Ana Montero-Castaño"
,"Anna Traveset"
,"Blanca Arroyo-Correa"
,"C. Heras"
,"Curro Molina"
,"C. O’Connor"
,"Concepción Ornosa"
,"C. Pérez-Iñigo Mora"
,"Carlos Lara-Romero"
,"T. Wood"
,"Concepción Ornosa"
,"Concepción Ornosa"
,"D. Paredes"
,"D. Romero"
,"David Alomar"
,"F. Javier Ortiz-Sánchez"
,"F. Javier Ortiz-Sánchez"
,"F. López-Núñez"
,"F. Perfectti"
,"Félix Torres"
,"F. Javier Ortiz-Sánchez"
,"Félix Torres"
,"F. Javier Ortiz-Sánchez"
,"H. Gaspar"
,"Ignasi Bartomeus"
,"Isabel María Liberal"
,"Isabel María Liberal"
,"J. Ascher"
,"J. Banaszak"
,"J. Bosch"
,"J. Dusmet"
,"J. Gonzalez-Varo"
,"J. L. Blanco-Pastor"
,"J. Loureiro"
,"J. R. Obeso"
,"J.A. González"
,"J. Bosch"
,"J.D. Alche"
,"J.D. Asís"
,"J.M. Gómez"
,"Javier Herrera"
,"Javier Morente-López"
,"Joana Costa"
,"José Iriondo"
,"Jose Luis Blanco-Pastor"
,"Juan Arroyo"
,"L. Castro"
,"Luis Navarro"
,"M. Campos"
,"M. Miñarro"
,"M. Porcel"
,"M.D. Martínez"
,"M.J. Dardón"
,"M.L. Fernández"
,"Montserrat Vilà"
,"Manuel Baena"
,"Marina Alonso Solano"
,"Mario Fernández-Mazuecos"
,"Carlos Martínez-Núñez"
,"Miguel A. González-Estévez"
,"Montserrat Vilà"
,"Pablo Vargas"
,"Pablo Vargas"
,"Pedro Coello"
,"Pilar de la Rúa"
,"R. Alcalá"
,"R. Carvalho"
,"R. Heleno"
,"P.J. Rey"
,"S. Castro"
,"S. Castro"
,"S.F. Gayubo"
,"Victoria Ferrero"
,"Volker Haeseler")

>>>>

```{r include=FALSE}
library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

master <- read.csv("../../Data/Processing_iberian_bees_raw/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
colnames(data)
colnames(master)
data[which(!data$accepted_name %in% master$accepted_name), ] # OK
master[which(!master$accepted_name %in% data$accepted_name), ]
# we are missing ~100 species, mostly parasites!
length(unique(data$accepted_name))
nrow(master)
```

@@

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# map
library(rgdal)
library(ggplot2)

@@

plotmap <- ggplot(data = data, aes(x = Longitude, y = Latitude)) + # plot the data
  geom_polygon(
    data = ESPTmap, aes(x = long, y = lat, group = group), # plot Spain
    color = "black", fill = "white"
  ) +
  coord_fixed() + # coord_fixed() ensures that one unit on the x-axis is the same length as one unit on the y-axis
  geom_point(color = "dodgerblue4", size = 2, shape = ".") + # graphical parameters for points
  stat_density2d(aes(
    x = jitter(Longitude), # create the density layer based on where the points are
    y = jitter(Latitude), fill = ..level.., alpha = ..level..
  ), # colour and transparency depend on density
  geom = "polygon", colour = "grey95", size = 0.3
  ) + # graphical parameters for the density layer
  scale_fill_gradient(low = "yellow", high = "red") + # set colour palette for density layer
  scale_alpha(range = c(.25, .5), guide = FALSE) + # set transparency for the density layer
  theme(
    axis.title.x = element_blank(), axis.text.x = element_blank(), # don't display x and y axes labels, titles and tickmarks
    axis.ticks.x = element_blank(), axis.title.y = element_blank(),
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    # text=element_text(size=18),legend.position = c(.9, .15),       # size of text and position of the legend
    panel.grid.major = element_blank(), # eliminates grid lines from background
    panel.background = element_blank()
  ) # set white background
# now plot, it takes a while!
plotmap
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width="70%"}
# code courtesy of Paco Rodriguez.
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
crs.geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs.laea <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

# Prepare data
data$Longitude <- as.numeric(data$Longitude)
data$Latitude <- as.numeric(data$Latitude)
data <- data %>% filter(!is.na(Latitude) & !is.na(Longitude))
# Keep points on balearic islands
data_1 <- data %>% filter(Province == "Balears, Illes")
data_1$Country_1 <- NA

# Filter points in Spain and Portugal
data$Country_1 <- NA
data$Country_1 <- maps::map.where(x = data$Longitude, y = data$Latitude)
data <- data %>% filter(Country_1 == "Spain" | Country_1 == "Portugal")
# Rbind points that are in Spain with the ones of Balearic islands
dat <- rbind(data, data_1)

datos.sf <- st_as_sf(dat,
  coords = c("Longitude", "Latitude"),
  crs = crs.geo
)

datos.laea <- st_transform(datos.sf,
  crs = crs.laea
)

@@

ras.counts <- rasterize(datos.laea[, "accepted_name"], ras,
  field = "accepted_name", fun = "count"
)
# str(datos.laea)
datos.laea$total <- datos.laea$Male + datos.laea$Female + datos.laea$Worker + datos.laea$Not.specified
ras.counts2 <- rasterize(datos.laea[, "total"], ras,
  field = "total", fun = "sum"
)
ras2 <- raster(datos.laea, resolution = 30000)
ras.richness <- rasterize(datos.laea[, "accepted_name"], ras2,
  field = "accepted_name", fun = function(x, ...) {
    length(unique(na.omit(x)))
  }
)
library(vegan)
ras.chao <- rasterize(datos.laea[, "accepted_name"], ras2,
  field = "accepted_name", fun = function(x, ...) {
    estimateR(t(as.matrix(x = table(x), nrow = 1, ncol = x)))[2, 1]
  }
) # 2 chao, 4 ACE #TO DO THIS WELL, we need to recover abundances.

# levelplot(log(ras.counts+1), margin = FALSE, scales = list(draw = FALSE),
#         par.settings = viridisTheme())
levelplot(log(ras.counts2 + 1),
  margin = FALSE, scales = list(draw = FALSE),
  par.settings = viridisTheme()
)
# levelplot(log(ras.richness+1), margin = FALSE, scales = list(draw = FALSE),
#         par.settings = viridisTheme())
# levelplot(ras.chao, margin = FALSE, scales = list(draw = FALSE),
#         par.settings = viridisTheme())
```

@@

```{r echo=FALSE, fig.align='center', out.width="70%"}
# Prepare top 20 records
top_20 <- data %>%
  filter(!accepted_name == "Apis mellifera") %>%
  group_by(accepted_name) %>%
  summarise(no_rows = length(accepted_name)) %>%
  arrange(-no_rows) %>%
  slice(1:20)
# Fix row position
top_20$accepted_name <- factor(top_20$accepted_name, levels = top_20$accepted_name)
# Plot
ggplot(top_20, aes(x = accepted_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8, face = "italic")) +
  ylab("Number of records") +
  xlab("Species") +
  scale_y_continuous(expand = c(0, 0))
```

@@

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Prepare top 20 records
by_year <- data %>% filter(!accepted_name == "Apis mellifera")

p1 <- ggplot(by_year, aes(Year)) +
  geom_histogram(fill = "#287D8EFF") +
  theme_classic() +
  ylab("Number of records") +
  scale_y_continuous(expand = c(0, 0))

@@

month_sum$month_name <- factor(month_sum$month_name, levels = c(
  "Jan", "Feb", "Mar",
  "Apr", "May", "Jun",
  "Jul", "Aug", "Sep",
  "Oct", "Nov", "Dec"
))

p2 <- ggplot(data = month_sum, aes(x = month_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  ylab("Number of records") +
  xlab("Months")

@@

p1 / p2
```

@@

# table of
# Name
# Contact
# Number of records as main dataholder
# Number of records as co-dataholder
# email?? Maually :(
# default <- par()$mar
# par(mar = c(13.1, 4.1, 4.1, 2.1))
# barplot(head(sort(table(data2$Authors.to.give.credit), decreasing = TRUE),20), las = 2, cex.names = 0.7)
# par(mar = default)
# Can be done prettier.
# Ideally, we can merge all authors, split and count individually.

@@

# sort(unique(author_list))[-1] #79 contributors!
# write.csv(sort(unique(author_list))[-1], "data/authors.csv")

@@

  "Alejandro Núñez",
  "Amparo Lázaro",
  "Ana Montero-Castaño",
  "Anna Traveset",
  "Blanca Arroyo-Correa",
  "C. Heras",
  "Curro Molina",
  "C. O’Connor",
  "Concepción Ornosa",
  "C. Pérez-Iñigo Mora",
  "Carlos Lara-Romero",
  "T. Wood",
  "Concepción Ornosa",
  "Concepción Ornosa",
  "D. Paredes",
  "D. Romero",
  "David Alomar",
  "F. Javier Ortiz-Sánchez",
  "F. Javier Ortiz-Sánchez",
  "F. López-Núñez",
  "F. Perfectti",
  "Félix Torres",
  "F. Javier Ortiz-Sánchez",
  "Félix Torres",
  "F. Javier Ortiz-Sánchez",
  "H. Gaspar",
  "Ignasi Bartomeus",
  "Isabel María Liberal",
  "Isabel María Liberal",
  "J. Ascher",
  "J. Banaszak",
  "J. Bosch",
  "J. Dusmet",
  "J. Gonzalez-Varo",
  "J. L. Blanco-Pastor",
  "J. Loureiro",
  "J. R. Obeso",
  "J.A. González",
  "J. Bosch",
  "J.D. Alche",
  "J.D. Asís",
  "J.M. Gómez",
  "Javier Herrera",
  "Javier Morente-López",
  "Joana Costa",
  "José Iriondo",
  "Jose Luis Blanco-Pastor",
  "Juan Arroyo",
  "L. Castro",
  "Luis Navarro",
  "M. Campos",
  "M. Miñarro",
  "M. Porcel",
  "M.D. Martínez",
  "M.J. Dardón",
  "M.L. Fernández",
  "Montserrat Vilà",
  "Manuel Baena",
  "Marina Alonso Solano",
  "Mario Fernández-Mazuecos",
  "Carlos Martínez-Núñez",
  "Miguel A. González-Estévez",
  "Montserrat Vilà",
  "Pablo Vargas",
  "Pablo Vargas",
  "Pedro Coello",
  "Pilar de la Rúa",
  "R. Alcalá",
  "R. Carvalho",
  "R. Heleno",
  "P.J. Rey",
  "S. Castro",
  "S. Castro",
  "S.F. Gayubo",
  "Victoria Ferrero",
  "Volker Haeseler"
)
